---
title: "Homework 5"
author: "Group 8"
date: "5/12/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ranger)
library(rpart)
library(tidymodels)
```

# Data and Setup

In this assignment, oriented towards cleaning data, we clean and analyze data on UK gas consumption and UK driver deaths.

First, we will download the data:
```{r, message=F}
ukg <- read_csv("https://codowd.com/bigdata/hw/hw6/ukgas.csv")
ukdd <- read_csv("https://codowd.com/bigdata/hw/hw6/ukdriverdeaths.csv")
```

# Clean Data
  
```{r}
# convert ukdd from wide to long
ukdd_long <- ukdd %>%
  pivot_longer(
    cols= -year,
    names_to = "month",
    values_to = "deaths")

# convert ukg from wide to long
ukg_long <- ukg %>%
  pivot_longer(
    cols= -year,
    names_to = "quarter",
    values_to = "gas") %>%
  mutate(
    # remove "Q" from quarter so numeric var
    quarter = as.numeric(str_sub(quarter, 2))
  )
```

```{r}
# define months in each quarter
q1_mos <- c("Jan", "Feb", "Mar")
q2_mos <- c("Apr", "May", "Jun")
q3_mos <- c("Jul", "Aug", "Sep")
q4_mos <- c("Oct", "Nov", "Dec")

ukdd_long <- ukdd_long %>%
  mutate(
    # create var indicating quarter of month
    quarter = as.numeric(case_when(
      month %in% q1_mos ~ 1, 
      month %in% q2_mos ~ 2,
      month %in% q3_mos ~ 3,
      month %in% q4_mos ~ 4
    ))
    )
```

```{r, message = FALSE}
# get total deaths in each year and quarter
ukdd_sum <- ukdd_long %>%
  group_by(year,quarter)%>%
  summarize(tot_deaths = sum(deaths)) %>%
  ungroup()
```

```{r}
uk_merge <- # merge quarterly datasets on year and quarter
  inner_join(ukdd_sum, ukg_long, 
             by=c("year","quarter"))%>%
  mutate(
    # create numeric var w/ year and quarter info
    time = year + (quarter-1)/4)
```

# Plot Results

Now that our data is clean, we will plot the relationship between quarterly deaths and quarterly gas consumption over time (on the same plot), and regress one against the other.


###  Plot quarterly UK deaths and UK gas consumption against time  

```{r}
uk_merge %>%
  ggplot()+
  geom_line(aes(x=time, y = tot_deaths, color = "Driver Deaths"))+
  geom_line(aes(x=time, y = gas, color = "Gas"))+
  labs(
    title = "UK Driver Deaths and Gas Consumption vs. Time",
    caption = "Source: codowd.com",
    x = "Time (Quarterly)",
    y =  "Count",
    color = NULL
  )
```
  
### Scatterplot of deaths against gas consumption

```{r, include = FALSE}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
uk_merge %>%
  ggplot(aes(x=gas, y = tot_deaths, color = as.factor(quarter)))+
  geom_point()+
  labs(title = "UK Driver Deaths vs. Gas Consumption",
    caption = "Source: codowd.com",
    x = "Time (Quarterly)",
    y =  "Count",
    color = "Quarter")+
  scale_color_manual(values = cbbPalette)
```

### Linear regression of deaths against gas consumption  

```{r, message = FALSE, warning = FALSE}
mod1 <- lm(tot_deaths ~ gas, data=uk_merge)

summary(mod1)
```
*Summary of Relationship*
  
The association between total driver deaths and gas consumption in the UK is surprisingly negative--an increase in gas consumption leads to fewer total driver deaths overall. However, this effect is somewhat small ($\beta=-.8323$), and not incredible significant ($p=0.0551$).  
